---
title: "Query Histocial Records from Columbia University History Lab - Declassification Engine API"
author: "Jianghanhan Li"
date: "2017-12-15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# R Package for History Lab Declassification Engine API 

Declassification Engine API is provided by [Columbia University History Lab](http://www.history-lab.org/). 

#### Setup

```{r eval=FALSE}
install.packages(c("devtools", "roxygen2", "testthat", "knitr"))
```

```{r eval=FALSE}
devtools::install_github(("hadley/devtools"), force = TRUE)
```

```{r}
library(devtools)
library(roxygen2)
library(testthat)
devtools::session_info()
```

```{r}
library(httr)
library(jsonlite)
library(rvest)
```

#### Dispaly Welcome Page `declass_welcome`

```{r}
declass_welcome <- function(return_type = "string"){
  url <- "http://api.declassification-engine.org/declass"
  message <- httr::content(GET(url))
  if (return_type == "html") {
    return(message)
  } else if (return_type == "string") {
    text <- rvest::html_text(message)
    return(text)
  } else {
    notice <- "Return_type not supported"
    return(notice)
  }
}

test_declass_welcome <- function() {
  print(declass_welcome("html"))  
  print(declass_welcome("string"))
  print(declass_welcome("xml"))
  print(declass_welcome())
}
test_declass_welcome()
```

#### Display all available document metadata fields `declass_fields`

```{r}
declass_fields <- function(return_type = "list") {
  url <- "http://api.declassification-engine.org/declass/v0.4/fields"
  if(return_type == "list") {
    parsed_json <- jsonlite::fromJSON(url)
    fields <- parsed_json$results$fields
    return(fields)
  } else if(return_type == "raw") {
    raw <- content(GET(url))
    return(raw)
  } else {
    notice <- "Return_type not supported"
    return(notice)
  }
}


test_declass_fields <- function() {
  print(declass_fields("list"))
  print(declass_fields("raw"))
  print(declass_fields())
}
test_declass_fields()
```

#### List all available collections `declass_collection`

```{r}
declass_collection <- function(return_type = "list"){
  url <- "http://api.declassification-engine.org/declass/v0.4/collections"
  parsed_json <- jsonlite::fromJSON(url)
  collections <- parsed_json$results$collections
  if (return_type == "list") {
    return(collections)
  } else if (return_type == "raw") {
    raw <- content(GET(url))
    return(raw)
  } else {
    notice <- "Return_type not supported"
    return(notice)
  }
}

test_declass_collection <- function() {
  print(declass_collection("list"))  
  print(declass_collection("raw"))
  print(declass_collection())
}

test_declass_collection()
```

#### Display the available entities for a particular collection `declass_collection_entity`

```{r}
declass_collection_entity <- function(collection) {
  if (collection %in% c("cpdoc", "clinton", "kissinger", "statedeptcables", "frus", "ddrs", "cabinet", "pdb")) {
    url <- paste("http://api.declassification-engine.org/declass/v0.4/entity_info/?collection=", collection, sep = "")
    parsed_json <- jsonlite::fromJSON(url)
    collection <- parsed_json$results$entity
    return(collection)
  } else {
    notice <- "Collection name and entities not found"
    return(notice)
  }
}

test_declass_collection_entity <- function() {
  print(declass_collection_entity("cpdoc"))  
  print(declass_collection_entity("clinton"))
  print(declass_collection_entity("kissinger"))
  print(declass_collection_entity("statedeptcables"))
  print(declass_collection_entity("frus"))
  print(declass_collection_entity("ddrs"))
  print(declass_collection_entity("cabinet"))
  print(declass_collection_entity("pdb"))
  print(declass_collection_entity(""))
}

test_declass_collection_entity()
```

#### Query data for a specified entity in a collection `declass_entity_data` 

```{r}
declass_entity_data <- function(collection, entity, page_size = 25, page = 1, num_pages = 10) {
  url <- paste("http://api.declassification-engine.org/declass/v0.4/entity_info/?collection=",collection, "&entity=", entity, "&page_size=", page_size, "&page=", page, sep = "")
  
   if (collection %in% c("cpdoc",  "kissinger", "cabinet") & entity %in% c("countries", "persons", "topics")) {
    }
   else if (collection %in% c("statedeptcables", "frus", "ddrs", "pdb") & entity %in% c("countries", "persons", "topics", "classifications")) {
    }
   else if (collection == "clinton" & entity %in% c("countries", "persons", "classifications")) {
      } 
  else {
    notice <- "Collection or entity name not found"
    return(notice)
    
  }
  print(url)
  parsed_json <- jsonlite::fromJSON(url)
  entity_data <- parsed_json$results
  num_pages <- num_pages - 1

  while (is.null(parsed_json$next_page) == FALSE & num_pages > 0) {
    print(parsed_json$next_page)
    parsed_json <- jsonlite::fromJSON(parsed_json$next_page)
    entity_data <- rbind(entity_data, parsed_json$results)
    num_pages <- num_pages - 1
  }
  return(entity_data)
}

test_declass_entity_data <- function() {
print(declass_entity_data("kissinger", "topics", page_size = 1, num_pages = 1))
print(declass_entity_data("kissinger", "topics", page_size = 1, num_pages = 2))
print(declass_entity_data("kissinger", "topics", page_size = 2, num_pages = 1))
print(declass_entity_data("kissinger", "topics"))
print(declass_entity_data("clinton", "topics"))
print(declass_entity_data("kissinger", "topics", page_size = 50, num_pages = 1))
}

test_declass_entity_data()
```

Here the topics in these collections were re-engineered using Topic Modeling, as exactly the same as available in the Declassification Engine API. 

#### Return random document ids with a limit `declass_query_random_ids`

```{r}
declass_query_random_ids <- function(limit = 10) {
  base_url <- "http://api.declassification-engine.org/declass/v0.4/random/?limit="
  if (limit <= 1024) {
    url <- paste(base_url, limit, sep = "")
    ids <- jsonlite::fromJSON(url)$results
    return(ids)
  } else {
    notice <- "Limit exceeds the maximum 1024"
    return(notice)
  }
}

test_declass_query_random_ids <- function() {
print(declass_query_random_ids(32))
print(declass_query_random_ids(1100))
}

test_declass_query_random_ids()
```

For those who don't know what they should be looking for here, we offer them randomly sampled ids. So they could check out the documents by ID later. 

#### Query document by ID `declass_query_id_data`

Allow user to manually input individual id or list of ids

```{r}
declass_query_id_data <- function(id_list) {
  base_url <- "http://api.declassification-engine.org/declass/v0.4/?ids="
  url <- paste(base_url,id_list, sep = "")
  doc <- jsonlite::fromJSON(url)
  return(doc)
}

# test 
id <- c("Clinton-74931")
declass_query_id_data(id)

```

#### Query documents based on randomly sampled ids `declass_query_random_id_data`

```{r}
declass_query_random_ids_data <- function(limit) {
  ids <- declass_query_random_ids(limit)
  id_list <- paste(ids$id, collapse = ",")
  id_list <- URLencode(id_list)
  return (declass_query_id_data(id_list))
}

# test
declass_query_random_ids_data(5)
```

